Improving cropland maps through tight integration of human and machine intelligence

### 2020-07-24 inst/manuscript/manuscript.Rmd
Substantial updates to analyses and organization. Manuscript, figures notebook, and analyses moved to inst/manuscript. Added scripts to data-raw/ for processing inputs into figures, headed to instances, etc (much of this is moving code out of analyses* notebooks).

Accuracy may also suffer due to error-inducing factors that are becoming somewhat more pronounced as a consequence of these technology advances, particularly with respect to algorithms. Advances in machine learning are helping to greatly improve classification skill, but these algorithms generally require large training datasets [@MaxwellImplementationmachinelearningclassification2018], particularly neural network-based "deep-learning" methods (cite). To satisfy this need for more training (and reference) samples, map-makers increasingly rely on visual interpretation of high resolution satellite or aerial imagery to collect training (or validation) samples [e.g. @XiongNominal30mCropland2017; @ChenGloballandcover2015; @StehmanKeyissuesrigorous2019]. Several web-based platforms have been developed to facilitate such efforts, which provide convenient and highly scalable tools for training data collection [e.g. @BeyCollectEarthLand2016; @Estesplatformcrowdsourcingcreation2016; @FritzGeoWikionlineplatform2012]. Visually interpreted training labels present two particular problems. The first is that such labels have inevitable interpretation errors that can vary substantially according to the skill of the interpreter [@Estesplatformcrowdsourcingcreation2016; @WaldnerConflationexpertcrowd2019]. These errors are typically not accounted for in reported accuracy metrics (cite Arthur's paper), thus the degree to which such labelling error impacts map accuracy is unclear. However, a study of how land cover map error propagates through subsequent analyses [@Esteslargeareaspatiallycontinuous2018] suggests training data quality could have a large impact. The second problem is that visual interpretation depends on high resolution imagery (<5 m), as lower resolutions make it difficult for a human analyst to discern cropland. Typically the only practical source for such imagery are "virtual globe" basemaps provided by Microsoft and Google, which are composed of mosaics of various high resolution satellite and aerial image sources that typically span 3-5 years of time within a single country [@LesivCharacterizingspatialtemporal2018]. This within-mosaic temporal variation can set up a temporal mismatch between the imagery being interpreted and the imagery being classified, which are usually from a different source (e.g. Landsat, Sentinel; @XiongNominal30mCropland2017) and often represent a different year. If a land change occurs in the interval between the two image sets (e.g. a new field was created), this can introduce error into the training data that is then passed on to the classifier. This source of error may be elevated in smallholder-dominated systems, where swidden practices are common (cite), or in rapidly developing agricultural frontiers [@ZengHighlandcroplandexpansion2018].  
<!-- - Why cropland mapping is hard -->
<!--     - 3 problems -->
<!--         - The target itself and its spectral variability and the spectral variability of its background -->
<!--             - Spectral and spatial resolution of sensors -->
<!--             - Poses a challenge for classification algorithms, which want both high resolution to be smaller than characteristic field size, but temporal contrast also -->

@@

# Methods
## System overview 
The mapping system is centered around four major components: 1) an image acquisition and pre-processing component, 2) a training data collection platform, 3) a machine learning component, and a 4) segmentation component. The first component is applied as a once-off step to generate an image catalog necessary for the mapping geography, while the second component initiates the the interactive image training and classification (component 3) pipeline that results in a pixel-wise posterior probability map of cropland probabilities on its final iteration (Fig. 1). The final segmentation step (component 4) is then initiated and applied to both the pre-processed primary images and the posterior probability maps, resulting in the final vectorized fields boundary maps. Each system component comprises an individual module software hosted on a GitHub repository (see data and software availability section for details).

## System components
### Imagery
**[Note: This will require updating to reflect changes to compositing methods]**
The primary image source used in our system is PlanetScope Analytic surface reflectance product [@PlanetTeamPlanetApplicationProgram2017], although can be readily adapted to apply to Sentinel and Landsat images. PlanetScope provides three visual (red, green, blue) and a near-infrared band at 3-4 m resolution at nominal daily frequency. Although these images are substantially pre-preprocessed and corrected for atmospheric effects, there are substantial residual errors from inter-sensor differences and the radiometric normalization process [@HouborgDailyRetrievalNDVI2018], substantial variation in the orientation of scene footprints, as well as a high frequency of cloud cover over the study region [@WilsonRemotelySensedHighResolution2016]. To correct for these factors, we developed a procedure for creating temporal composites representing the primary growing and non-growing seasons within a single year.

@@

PlanetScope imagery is accessed via the Planet API, and an initial order is placed wherein the imagery representing all available dates in the two compositing periods and within a bounding box covering the mapping geography is collected and transferred directly to the system's cloud storage (hosted on Amazon Web Services S3).

We then transform all images into an "analysis ready data" (ARD) product [@DwyerAnalysisReadyData2018] that is prepared against a tiling grid that is situated within a larger 1 degree resolution grid that covers the entire continent. Each 1 degree cell defines the minimum mapping area of interest (Fig. 2A) is divided into 400 0.05 degree cells (Fig. 2B), which provides the scale for creating image composite tiles. The ARD process simply clips all available imagery intersecting each 0.05 cell to the extent of that cell, regardless of cloud cover.

@@

```{r, echo = FALSE, out.width="100%", fig.cap="An overview of analytical scales used in the mapping platform, including the 1 degree areas of interest defining the minimum mapping geography (A), the 0.05 degree grid used for creating image composite tiles (B), and the 0.005 degree resolution reference grid used for collecting training data and distributed computing (C).", fig.align='center'}
knitr::include_graphics('figures/figure2.png')
```
The temporal compositing process is applied to the ARD imagery for two time periods, the dry season period (December-February) and the entire wet season (May - September). The longer period is necessary for the wet season because of the high frequency of cloud cover, which severely limits the number of clear scenes over any given area. For each pixel in each image in each ARD temporal stack, we calculate two weights, with the first being:

@@

Once these two weights are calculated, the final composited pixel value for each of the four PlanetScope bands is calculated as:

@@

Which is a weighted mean for each pixel for each band *B* for the particular compositing period. The composited tiles are then added to the S3 store (Fig. 1), where they are stored as cloud-optimized geotiffs, and a "slippy map" rendering is created for each composite using RasterFoundry (cite). This web-based rendering of the image is used in the training data platform (see section X).

### Labelling platform
Both training and reference data are collected by a custom crowdsourcing platform, which we refer to here for simplicity's sake as *labeller*. The original version of *labeller* was designed to work with Amazon Web Services' Mechanical Turk job marketplace [@Estesplatformcrowdsourcingcreation2016]. The basic structure of the system remains the same, but we have re-engineered it to be a standalone platform that 1) allows us to register and pay workers (the human supervisors) directly to conduct digitizing work, and 2) drives and assesses the overall machine learning process. The system runs on a Linux Virtual machine hosted on an AWS EC2 instance and is comprised of a database (PostGIS/Postgres), a mapping interface (OpenLayers 3), an image server (RasterFoundry), and a set of utilities for managing mapping and assessing assignments and converting them to rasterized labels for the machine learning process. Each *labeller* instance is focused on a specific mapping geography (e.g. one to several 1 degree cells in Fig. 2A), thus several instances are set up to run in parallel to cover a larger mapping geography such as Ghana.

Here we provide an overview of *labeller*'s architecture within the context of the labeling workflow within a single instance.

#### Mapping workflow
##### Selecting training and reference sites
The labelling process begins with a python script that randomly selects a subset (e.g. 100) of the 0.005 degree grid (Fig. 2C) cells that intersect the *labeller* instance's specific area of interest (e.g. shaded grey cell in Fig. 2A). These site names are stored in a Postgres table that represents the mapping queue. Another python process converts each site into a mapping *task* that has a specified number of *assignments* (maps drawn by an individual worker) that must be completed before the task is complete.

##### Mapping assignments
Workers registered in the system log in to the mapping platform (built with Flask) and navigate to the OpenLayers-based field mapping interface (Fig. 3). There the worker is presented with a white target box representing the 0.005 degree cell that was randomly selected to label, a set of digitizing tools, and different image backdrops, which include false and true color renderings of both the growing season and off-growing season PlanetScope images covering that cell, as well as a basemap provided by Bing maps. The worker uses the polygon drawing tool to digitize the boundaries in crop fields visible within the PlanetScope overlays that intersect the grid target, following a set of pre-defined digitizing rules. To aid with interpretation, the worker toggles between the various seasonal renderings and the satellite base map to help form a judgement about what constitutes a field within the PlanetScope imagery. The worker assigns each completed polygon a label that describes the field class (e.g. annual cropland, tree crop, fallow), and after completing all visible fields saves the full set of polygons to the server. The worker is then automatically  taken to their next assignment at a different randomly selected site.

@@

##### Processing completed assignments
Upon submission, the workers' polygons, if there are any--the worker will submit the assignment without mapping if they see no fields in the image--are cleaned to fix any topological irregularities that arose during digitization (see supporting information (SI)) and stored in a PostGIS table. The assignment that was just completed represents one of two possible types of tasks: 1) an accuracy task; 2) a labeling task for training machine learning. If is was the former type, an `R` script is called that executes a series of map comparisons between the worker's polygons and the reference polygons, assigning a score based on the following formula:

@@

If the worker's completed assignment was a labelling task, their maps remain stored in the database until the remaining assignments needed to complete the task are submitted by other workers. Once complete, an R-based consensus labelling routine is invoked, which combines the maps from all workers who undertook the task into a single consensus label using a Bayesian merging approach:

@@

Where $\theta$ represents the true cover type of a pixel (field or not field), *D* is the worker's label of that field, and $W_i$ is an individual worker.  P($\theta$|D) is therefore the probability that the actual cover type is what the workers who labelled it says it is, while P(W$_i$|D) is the worker's average score over the accuracy assessment assignments they have mapped to date, and P(W$\theta$|D, *W$_i$*) is the worker's label for that pixel. This approach therefore uses the overall accuracy of each worker to weight their labels as they are combined with other workers' labels for a given pixel (see SI for further details of equation).

Once a the full set of sites has been completed, they are randomly partitioned into training and validation sets according to predetermined proportions.

@@

### Classification pipeline
Upon completion of the labels, *labeller* automatically launches an ephemeral Elastic Map Reduce cluster consisting of tens to hundreds of instances, depending on the size of the mapping geography. We identifed an optimal cluster of 50-60 instances (m4.2xlarge) each having 8 virtual CPUs and 30 GB RAM, which ensured minimal idle time for any single node.

#### Feature extraction
The first step is the extraction of additional features from the growing season/non-growing season PlanetScope imagery. Previous work showed that a large number of simple features that summarize the statistical properties of reflectance and vegetatiation indices in local neighborhoods are highly effective for classifying smallholder croplands [@Debatsgeneralizedcomputervision2016]. We followed this logic in this study, but opted for a smaller feature space to accomodate the storage and memory requirements of a mapping geography orders of magnitude larger than that initial work. Here, we used a set of 24 features, comprising the eight original composite bands (four from each season), an 11X11 mean of each band, as well as the standard deviation of each band within a 5X5 window (Table X).

\begin{center}Table X. List of features.\end{center}

@@

Feature extraction and the conversion of image features is handled by a combination of `GeoTrellis`, `GeoPySpark`, `rasterio`, and `RasterFrames`. These collectively extract subsets of imagery from the PlanetScope temporal composites, derive the features (Table X), and convert these into `RasterFrames`, which are an Apache Spark compatible data object that is designed specifically for raster data. Features are extracted on the fly for each cell in the training and validation sets, a functionality enabled by storing the image composites as Cloud-optimized Geotiffs (COGs), which allows fast windowed reads of image subsets.

@@

### The active learning loop
After fitting and model evaluation a second prediction process is undertaken that initiates the active learning process. The feature extraction process is repeated for the rest of the mapping geography that falls outside of the training and validation sets, but applied to a subset of randomly drawn pixels from each cell in order to reduce computation demand. The fitted model is then applied to these features in order to calculate the posterior probability of cropland presence, and an uncertainty criterion [@DebatsIntegratingactivelearning2017] is calculated for each grid cell:

@@

In the second step, a Sobel filter is then applied to the green, red, and near-infrared mean-shifted bands [**I assume**], and weighted edge image is computed using weights of 1, 1, and 2 to each band for each season [**for both growing season on and off?**]. A watershed algorithm (**cite**) is then run on the weighted edge image, with a high level of segmentation specified (**2400** segements per tile).

@@

In the fourth step, the merged polygons are overlaid with the posterior probabilities resulting from the final active learning loop, and polygons in which the average posterior probability is greater than a predetermined  threshold (typically 0.5, but is possible to vary locally) are retained as field polygons.

@@

### Accuracy assessment
**[Put in somewhere above text about training reference samples]**
The accuracy assessment procedure described in section 2.2 is used strictly to assess performance gains in the active learning loop, according to suggested best practice guidelines [@ElmesAccountingtrainingdata2019]. To assess the accuracy of final map products, following these same guidelines we used an independent map reference sample. To develop this sample, we first made a random selection of grid cells within each mapping AOI, with the selected number scaled relative to the AOI. Next, a primary supervisor examined the image composites within each randomly selected cell and placed three rectangular polygons of ~0.1 ha area in areas that were fully contained within in one of three classes: cropland, non-cropland, potentially cropland or non-cropland. In cases where less than three classes were present within the cell, the supervisor placed polygons in proportion to the dominant class. For example, in the case where the entire cell was taken up by non-cropland, the supervisor placed all three polygons in non-cropland. If two-thirds of the cell was non-cropland, and one-third ambiguous, two polygons would be placed in non-cropland, and one in the ambiguous class. After placing these polygons, the supervisor removed their classes from the polygons and sent them to three separate experts, who labelled the polygons according to their interpretation of the underlying imagery. The four sets of labels were then combined and the overall agreement between them was calculated. This agreement provided the upper bound on knowable map accuracy.

## Application and evaluation of the system
[Description of map creation process]
To create a map of Ghana, we divided the country into 16 distinct mapping zones, or areas of interest (AOIs), which were developed by grouping together each 1X1 degree cell (Figure 2A) fully contained with the tiles (Figure 2B) belonging to any adjacent 1X1 cells that overlapped with neighboring countries. The resulting AOIs ranged from 12,160 to 23,535 km$^2$ in extent (average = 15,457 km$^2$), and each was used to run a separate active learning process and to create a final segmented map.

Prior to initiating active learning in each AOI, we grouped the AOIs into three zones: a northern zone comprising the 6 northernmost AOIs, a central to southeastern zone comprising the 3 middle and 3 southeastern AOIs, and a southwestern zone made up of the AOIs falling in the forest zone (see Figure SX in SI). Within each zone group, we randomly selected 500 points for initial labelling by the mapping team, which were used to train the first iteration of each RandomForest model run for the zone group. At the outset of mapping in each AOI, the mapping team labelled 100 randomly selected points within each AOI that were set aside for model validation, and subsequent iterations targeted new training sites for collection in each AOI based on classification uncertainty.

To more fully evaluate the capacities of different components of the system, we performed several experiments. In the first experiments, we evaluated the degree to which training label quality impacts final map accuracy, by comparing maps created by training the RandomForest algorithm with labels created by the least accurate worker, the most accurate worker, in addition to those trained with the consensus labels. This test aligns with Tier 1 standards for accounting for the impact of training data error on final map accuracy, as recommended by @ElmesAccountingtrainingdata2019. Next, we evaluated the degree to which training sample selection via active learning impacted map accuracy relative to random selection of additional training sites.

@@

@@

## Image catalog

We generated a complete image catalog for the `r nrow(tiles)` tiles intersecting Ghana, spanning `r km2` km$^2$. This entailed processing all PlanetScope imagery intersecting the total extent of these tiles for the periods May-September, 2018 (the growing season) and December, 2018 to February, 2019 (the subsequent dry season). For cloudier regions (AOIs 10, 11, 13, 14, 16) we started the dry season window in November. The compositing processes was applied to render the resulting seasonally contrasting image pairs for each tile (**Figure Xa**).

@@

## The active learning process
**[Fill in with intro]**. The average number of iterations per active learning process was **4**, as overall performance gains, as assessed against validation sets, saturated quickly (see **Figure X** in SI).

@@

>>>>

Accuracy may also suffer due to error-inducing factors that are becoming somewhat more pronounced as a consequence of these technology advances, particularly with respect to algorithms. Advances in machine learning are helping to greatly improve classification skill, but these algorithms generally require large training datasets [@MaxwellImplementationmachinelearningclassification2018], particularly neural network-based "deep-learning" methods (cite). To satisfy this need for more training (and reference) samples, map-makers increasingly rely on visual interpretation of high resolution satellite or aerial imagery to collect training (or validation) samples [e.g. @XiongNominal30mCropland2017; @ChenGloballandcover2015; @StehmanKeyissuesrigorous2019]. Several web-based platforms have been developed to facilitate such efforts, which provide convenient and highly scalable tools for training data collection [e.g. @BeyCollectEarthLand2016; @Estesplatformcrowdsourcingcreation2016; @FritzGeoWikionlineplatform2012]. Visually interpreted training labels present two particular problems. The first is that such labels have inevitable interpretation errors that can vary substantially according to the skill of the interpreter [@Estesplatformcrowdsourcingcreation2016; @WaldnerConflationexpertcrowd2019]. These errors are typically not accounted for in reported accuracy metrics, despite the fact that they can introduce substantial error into the resulting maps and subsequent analyses [@ElmesAccountingtrainingdata2020; @Esteslargeareaspatiallycontinuous2018]. The second problem is that visual interpretation depends on high resolution imagery (<5 m), as lower resolutions make it difficult for a human analyst to discern cropland. Typically the only practical source for such imagery are "virtual globe" basemaps provided by Microsoft and Google, which are composed of mosaics of various high resolution satellite and aerial image sources that typically span 3-5 years of time within a single country [@LesivCharacterizingspatialtemporal2018]. This within-mosaic temporal variation can set up a temporal mismatch between the imagery being interpreted and the imagery being classified, which are usually from a different source (e.g. Landsat, Sentinel; @XiongNominal30mCropland2017) and often represent a different year. If a land change occurs in the interval between the two image sets (e.g. a new field was created), this can introduce error into the training data that is then passed on to the classifier. This source of error may be elevated in smallholder-dominated systems, where swidden practices are common (cite), or in rapidly developing agricultural frontiers [@ZengHighlandcroplandexpansion2018].  
<!-- - Why cropland mapping is hard -->
<!--     - 3 problems -->
<!--         - The target itself and its spectral variability and the spectral variability of its background -->
<!--             - Spectral and spatial resolution of sensors -->
<!--             - Poses a challenge for classification algorithms, which want both high resolution to be smaller than characteristic field size, but temporal contrast also -->

@@

# Methods
## System overview 
The mapping system is centered around four components: 1) an image acquisition and compositing component, 2) a training data collection platform, 3) a machine learning component, and 4) a segmentation component. The first component is applied as a once-off step to generate an image catalog covering the mapping geography, while the second component initiates an interactive image training and classification (component 3) pipeline that results in a pixel-wise posterior probability map of cropland probabilities on its final iteration (Fig. 1). The final segmentation step (component 4) is then initiated and applied to both the pre-processed primary images and the posterior probability maps, resulting in vectorized field boundaries. Each system component comprises an individual software module designed to be implemented on cloud computing architecture, and available on a GitHub repository (see data and software availability section for details).

## System components
### Imagery
**[Note: This will require updating to reflect changes to compositing methods]**
The primary image source used in our system is the PlanetScope Analytic surface reflectance product [@PlanetTeamPlanetApplicationProgram2017], although it can be readily adapted to apply to other images (e.g. Sentinel and Landsat). PlanetScope provides three visual (red, green, blue) and a near-infrared band at 3-4 m resolution at nominal daily frequency. Although these images are already pre-preprocessed and corrected for atmospheric effects, there are residual errors from inter-sensor differences and the radiometric normalization process [@HouborgDailyRetrievalNDVI2018], variation in the orientation of scene footprints, as well as a high frequency of cloud cover over the study region [@WilsonRemotelySensedHighResolution2016]. To correct for these factors, we developed a procedure for creating temporal composites representing the primary growing and non-growing seasons within a single year.

@@

PlanetScope imagery is accessed via the Planet API, and an initial order is placed wherein the imagery representing all available dates in the two compositing periods and within a bounding box covering the mapping geography is collected and transferred directly to the system's cloud storage (Amazon Web Services S3).

We then transform all images into an "analysis ready data" (ARD) product [@DwyerAnalysisReadyData2018] that is prepared against a tiling grid that is situated within a larger 1 degree resolution grid that covers the entire continent. Each 1 degree cell defines the minimum mapping area of interest (Fig. 2A) is divided into 400 0.05 degree cells (Fig. 2B), which provides the scale for creating image composite tiles. The ARD process simply clips all available imagery to the extent of each 0.05 cell it intersects, regardless of cloud cover.

@@

```{r, echo = FALSE, out.width="100%", fig.cap="An overview of analytical scales used in the mapping platform, including A) the areas of interest defining the minimum mapping geography (solid black lines; dotted lines indicate boundaries of 1 degree grid), the B) 0.05 degree grid used for creating image composite tiles, and the C) 0.005 degree resolution reference grid used for collecting training data and distributed computing.", fig.align='center'}
knitr::include_graphics('figures/figure2.png')
```
The temporal compositing process is applied to the ARD imagery for the time period of interest, which in the case of this study is one of two multi-month seasons, the primary growing and dry seasons for a single agricultural year. Imagery from two seasons helps to improve the performance of cropland classifiers [@Debatsgeneralizedcomputervision2016], while having the seasons in same year helps to minimize differences caused by land change. For each pixel in each image in each ARD temporal stack for a given season, we calculate two weights, with the first being:

@@

Once these two weights were calculated, the final composited pixel value for each of the four PlanetScope bands was calculated as:

@@

Which is a weighted mean for each pixel for each band *B* for the particular compositing period. The composited tiles were then added to the S3 store (Fig. 1), where they are stored as cloud-optimized geotiffs, and a "slippy map" rendering is created for each composite using RasterFoundry (cite). This web-based rendering of the image is used in the training data platform (see **section X**).

### Labelling platform
Both training and reference data were collected by a custom crowdsourcing platform, which we refer to here for simplicity's sake as *labeller*. The original version of *labeller* was designed to work with Amazon Web Services' Mechanical Turk job marketplace [@Estesplatformcrowdsourcingcreation2016]. The basic structure of the system remains the same, but we have re-engineered it into a standalone platform that 1) allows us to enroll and pay workers (the human supervisors) directly to conduct digitizing work, and 2) drives and assesses the overall machine learning process. The system runs on a Linux Virtual machine hosted on an AWS EC2 instance and is comprised of a database (PostGIS/Postgres), a mapping interface (OpenLayers 3), an image server (RasterFoundry), and a set of utilities for managing mapping and assessing assignments and converting them to rasterized labels for the machine learning process. Each *labeller* instance is focused on a specific mapping geography (e.g. one to several 1 degree cells in Fig. 2A)

Here we provide an overview of *labeller*'s architecture.

#### Mapping workflow
##### Selecting training and reference sites
The labelling process begins with a python script that randomly selects a subset (e.g. 100) of the 0.005 degree grid (Fig. 2C) cells that intersect the *labeller* instance's specific area of interest (e.g. shaded grey cell in Fig. 2A). These site names are stored in a Postgres table with the PostGIS extension that represents the mapping queue. Another python process converts each site into a mapping *task* that has a specified number of *assignments* (maps drawn by an individual worker) that must be completed before the task is complete.

##### Mapping assignments
Workers registered in the system log in to the mapping platform (built with Flask) and navigate to the OpenLayers-based field mapping interface (Fig. 3). There the worker is presented with a white target box representing the 0.005 degree cell that was randomly selected to label, a set of digitizing tools, and different image backdrops, which include false and true color "slippy map" renderings (generated by Raster Foundry) of both the growing season and off-growing season PlanetScope images covering that cell, as well as basemap imagery provided by Bing and ESRI. The worker uses the polygon drawing tool to digitize the boundaries in crop fields visible within the PlanetScope overlays that intersect the grid target, following a set of pre-defined digitizing rules. To aid with interpretation, the worker toggles between the various seasonal renderings and the satellite base map to help form a judgement about what constitutes a field within the PlanetScope imagery. The worker assigns each completed polygon a label that describes the field class (e.g. annual cropland, tree crop, fallow), and after completing all visible fields saves the full set of polygons to the server. The worker is then automatically taken to their next assignment at a different randomly selected site.

@@

##### Processing completed assignments
Upon submission, the workers' polygons, if there are any--the worker will submit the assignment without mapping if they see no fields in the image--are cleaned to fix any topological irregularities that arose during digitization (see supporting information (SI)) and stored in a PostGIS table. The assignment that was just completed represents one of two possible types of tasks: 1) an accuracy task; 2) a labeling task for training the machine learning algorithm. If it was the former type, an `R` script is called that executes a series of map comparisons between the worker's polygons and the reference polygons, assigning a score based on the following formula:

@@

If the worker's completed assignment was a labelling task, their maps remain stored in the database until the remaining assignments needed to complete the task are submitted by other workers. Once complete, a `R`-based routine is invoked, which combines the maps from all workers who undertook the task into a single consensus label using a Bayesian merging approach:

@@

Where $\theta$ represents the true cover type of a pixel (field or not field), *D* is the worker's label of that field, and $W_i$ is an individual worker.  P($\theta$|D) is therefore the probability that the actual cover type is what the workers who labelled it says it is, while P(W$_i$|D) is the worker's average score over the accuracy assessment assignments they have mapped to date, and P(W$\theta$|D, *W$_i$*) is the worker's label for that pixel. This approach therefore uses the overall accuracy of each worker to weight their labels as they are combined with other workers' labels for a given pixel (see SI for further details). The Bayesian risk of the completed final is then assessed, which measures the average level of consensus for the label assigned to each pixel in a given site, with values closer to 0 indicating complete consensus, and thus the highest confidence, while values approaching 1 reveal progressively increasing disagreement (see SI).

Once all training sites are completed, they are randomly partitioned into training and validation sets according to predetermined proportions.

@@

### Classification pipeline
Upon completion of the labels, *labeller* automatically launches an ephemeral Elastic Map Reduce cluster consisting of tens of instances, depending on the size of the mapping geography.

#### Feature extraction
The first step is the extraction of additional features from the growing season/non-growing season PlanetScope imagery. Previous work showed that a large number of simple features that summarize the statistical properties of reflectance and vegetatiation indices in local neighborhoods are highly effective for classifying smallholder croplands [@Debatsgeneralizedcomputervision2016]. We followed this logic in this study, but opted for a smaller feature space to accomodate the storage and memory requirements of mapping geography that were several orders of magnitude larger. Here, we used a set of 24 features, comprising the eight original composite bands (four from each season), an 11X11 mean of each band, as well as the standard deviation of each band within a 5X5 window (Table 1).

\begin{center}Table 1. List of features.\end{center}

@@

Feature extraction and the conversion of image features is handled by a combination of `GeoTrellis`, `GeoPySpark`, `rasterio`, and `RasterFrames`. These collectively extract subsets of imagery from the PlanetScope temporal composites, derive the features (Table 1), and convert these into `RasterFrames`, which are an Apache Spark compatible data object that is designed specifically for raster data. Features are extracted on the fly for each cell in the training and validation sets, a functionality enabled by storing the image composites as Cloud-optimized Geotiffs (COGs), which allows fast windowed reads of image subsets.

@@

### The active learning loop
After fitting and model evaluation a second prediction process is undertaken that initiates the active learning process. The feature extraction process is repeated for the rest of the mapping geography that falls outside of the training and validation sets, but applied to a subset of randomly drawn pixels from each cell in order to reduce computational demand. The fitted model is then applied to these features in order to calculate the posterior probability of cropland presence, and an uncertainty criterion [@DebatsIntegratingactivelearning2017] is calculated for each grid cell:

@@

In the second step, a Sobel filter is then applied to the green, red, and near-infrared mean-shifted bands [**I assume**], and weighted edge image is computed using weights of 1, 1, and 2 to each band for the dry season only [**or growing season?**]. A watershed algorithm (**cite**) is then run on the weighted edge image, with a high level of segmentation specified. In this case, we specified **2400** segments per tile.

@@

In the fourth step, the merged polygons are overlaid with the posterior probabilities resulting from the final active learning loop, and polygons in which the average posterior probability is greater than a predetermined  threshold (typically 0.5, but can vary locally) are retained as field polygons.

@@

## Application and evaluation of the system
To create a map of Ghana, we divided the country into 16 distinct mapping zones, or areas of interest (AOIs), which were developed by grouping together each one degree cell (Figure 2A) fully contained with the tiles (Figure 2B) belonging to any adjacent one degree cells that overlapped with neighboring countries. The resulting AOIs ranged from 12,160 to 23,535 km$^2$ in extent (average = 15,457 km$^2$), and each was used to run a separate active learning process and to create a final segmented map.

Prior to initiating active learning in each AOI, we grouped them into three zones: a northern zone comprising the 6 northernmost AOIs, a central to southeastern zone comprising the 3 middle and 3 southeastern AOIs, and a southwestern zone made up of the AOIs falling in the forest zone (see Figure SX in SI). Within each zone group, we randomly selected 500 points for initial labelling by the mapping team, which were used to train the first iteration of each RandomForest model run for the zone group. At the outset of mapping in each AOI, the mapping team labelled 100 randomly selected points within each AOI that were set aside for model validation, and subsequent iterations targeted new training sites for collection in each AOI based on classification uncertainty.

To evaluate the results, we undertook a number of analyses.

### Accuracy assessment

The accuracy assessment procedure described in section 2.2 was used strictly to assess performance gains in the active learning loop, according to suggested  best practices [@ElmesAccountingtrainingdata2020]. To independently assess the categorical accuracy of final map products, we followed guidelines provided by [@StehmanKeyissuesrigorous2019] to create a map reference sample. To develop this sample, we first made a stratified random sample of grid cells across the entire country, using the mapped field boundaries to define the two strata. Next, a primary supervisor examined the image composites in each randomly selected cell, and placed a square polygon of ~0.1 ha on the primary class intersecting the centerpoint of the cell. The square was shifted so that it was contained entirety with the intersected class: cropland or non-cropland. Where the supervisor was uncertain about the class label, one of two additional classes were selected: likely cropland or likely non-cropland. After placing these polygons, the supervisor removed their classes from the polygons and sent them to three separate experts, who labelled the polygons according to their interpretation of the underlying imagery. The four sets of labels were then combined and the overall agreement between them was calculated. This agreement provided the upper bound on knowable map accuracy. See the SI for further details on the creation and analysis of the map reference sample.

To assess the accuracy of the segmented field boundaries, we compared the size class distributions of the segmented field boundaries against those of the  workers' digitized polygons at map validation sites. We calculated the differences in the measures of central tendency (mean, median, mode) between the two datasets within each validation site, and at the scale of the AOI.

To further evaluate the capacities of different components of the system, we performed several further experiments. In the first experiment, we evaluated the degree to which training label quality impacts final map accuracy, by comparing maps created by training the RandomForest algorithm with labels created by the least accurate worker, the most accurate worker, as well as those trained with the consensus labels. This test aligns with Tier 1 standards for accounting for the impact of training data error on final map accuracy [@ElmesAccountingtrainingdata2020]. Next, we evaluated the degree to which training sample selection via active learning impacted map accuracy relative to random selection of additional training sites. Lastly, we repeated the entire map development and map assessment process for one AOI for the 2019 agricultural year, in order to assess the ability of the system to create annually updated maps and detect inter-annual change.

@@

## The characteristics of Ghanaian cropland
Using the final mapped results, we calculated the estimated area of arable cropland in Ghana, as well as the characteristics of field sizes through the country. We further used the uncertainty information encoded in the cropland probability images produced by the Random Forest classifier to provide further insight into the characteristics of cropland active-fallow cycles.

@@

## Image catalog
We generated a complete image catalog for the `r nrow(tiles)` tiles intersecting Ghana, spanning `r km2` km$^2$. This entailed processing all PlanetScope imagery intersecting the total extent of these tiles for the periods May-September, 2018 (the growing season) and December, 2018 to February, 2019 (the subsequent dry season). The longer period was necessary for the growing season because of the high frequency of cloud cover, which severely limits the number of clear scenes over any given area. For cloudier regions (AOIs 10, 11, 13, 14, 16) we started the dry season window in November. The compositing processes was applied to render the resulting seasonally contrasting image pairs for each tile (**Figure Xa**). We also collected and composited imagery for one of the 16 AOIs for the two seasons of the 2019 agricultural year. All images were converted into Raster Foundry slippy maps for presentation in *labeller*.

@@

## Active learning
### Training data collection
```{r, echo=FALSE, message=FALSE}
initial <- train_val_sites$initial %>% as_tibble %>% group_by(Cluster) %>%
  count()
active <- train_val_sites$active %>% as_tibble %>% filter(usage == "train") %>% 
  group_by(aoi) %>% count()
typicaln <- raster::modal(initial$n) + raster::modal(active$n)
aoi10n <- (initial %>% filter(Cluster %in% 3) %>% pull(n)) + 
  (active %>% filter(aoi == "labeller10") %>% pull(n))
aoi14n <- (initial %>% filter(Cluster %in% 3) %>% pull(n)) + 
  (active %>% filter(aoi == "labeller14") %>% pull(n))
aoi15n <- (initial %>% filter(Cluster %in% 2) %>% pull(n)) + 
  (active %>% filter(aoi == "labeller15") %>% pull(n))
aoi3init <- initial %>% filter(!Cluster %in% 1:3) %>% pull(n)
aoi3n <- aoi3init + (active %>% filter(aoi == "labeller3") %>% pull(n))

initial_total <- train_val_sites$initial %>% as_tibble() %>% distinct(name) %>% 
  count()
active_total <- sum(active$n)
validate_total <- train_val_sites$active %>% as_tibble() %>% 
  filter(usage == "validate") %>% distinct(name) %>% count()
```

After initial training, the active learning process was run for 3 iterations for 12 of the 16 AOIs, resulting in `r typicaln` labels per AOI (Figure S2A). AOIs 10 and 14 stopped after one and two iterations, respectively, as they started with high initial validation accuracies (>83%) and showed little subsequent improvement, thus were trained with `r aoi10n` -  `r aoi14n` samples. AOI 15 was run for 4 iterations and had a training sample size of `r aoi15n`, while AOI 3 underwent a second active learning cycle because the model produced during the first cycle was inaccurate (see SI). In this second run, `r aoi3init`  initial training sites randomly selected within the AOI were used (Figure S2A), followed by 2 subsequent active learning subsequent iterations, resulting in a training sample of `r aoi3n`. The total number of unique training and validation sites across the country were `r initial_total + active_total` and `r validate_total`, respectively.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
data("label_summary")  
tsites <- label_summary$nlabels %>% group_by(kml_type) %>% 
  summarize(tot = sum(n), mu = round(mean(n)), med = round(median(n)), 
            stdev = round(sd(n)), lower = min(n), upper = max(n))

# training rate
train_rate <- label_summary$nlabels %>% 
  tidyr::pivot_wider(names_from = kml_type, values_from = n) %>% 
  rename_all(funs(c("ID", "tr", "tref"))) %>% 
  mutate(trate = tref / tr)
```

The distribution of training and validation sample collection effort was divided across `r nrow(label_summary$score_stats)` labellers, with a core team of `r label_summary$nlabels %>% filter(kml_type == "Training" & n > 1000) %>% nrow()` who mapped more than 1,000 sites each. As each training site was mapped by 4 separate labellers, `r tsites$tot[1]` sets of vector labels were made. Each labeller digitized an average of `r tsites$mu[1]` (see Figure S3A for more details on labelling effort).

This accuracy of the labellers was assessed `r tsites$tot[2]` times against `r length(unique(label_summary$scores$name))` unique training reference sites (Figure S3A), with each labeller assessed an average of `r tsites$mu[2]` times at a rate of 1 training reference site for every `r round(1 / (tsites$mu[2] / tsites$mu[1]), 2)` training site mapped. The average of each labeller's average accuracy score was `r round(mean(label_summary$score_stats$Mean), 2)` (range `r round(min(label_summary$score_stats$Mean), 2)` to `r round(max(label_summary$score_stats$Mean), 2)`; see Figure S3B for detailed score distributions).

After each site was mapped by four labellers, consensus labels were generated.

### Model performance

**[Fill in with intro]**. The average number of iterations per active learning process was **4**, as overall performance gains, as assessed against validation sets, saturated quickly (see **Figure X** in SI).

@@

###
